Viscous Transport in Eroding Porous Media

### 2019-10-02 docs/transport/erosion_transport.tex
finished up nick's comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In equations~\eqref{eqn:velocityDLP} and~\eqref{eqn:deformationDLP},
$\uu$ and its deformation tensor are written as layer potentials in
$\RR^2$, but the quadrature method we introduce in
Section~\ref{sec:method} requires a complex-valued representation of the
layer potentials. The first step to form a complex representation is to
write the Laplace double-layer potential as the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Re (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \nick{ \CC(\Omega) },
  \label{eqn:laplaceComplex}
\end{align}
where $x,y \in \CC$ are the complex counterparts of $\xx,\yy \in \RR^2$,
and $\eta$ the complex counterpart of $\eeta$. \nick{We use the notation $\CC(\Omega)$ to denote the physical domain $\Omega$ represented in complex coordinates.}
\todo[inline]{What do you think of this notation I introduced to distinguish between $\Omega$ as a real versus complex domain? I was skimming over this section and was very confused about what $x$ and $y$ were, having naively assumed they were physical $x$ and $y$ coordinates instead of complex numbers. Hence I wanted to make their definitions more prominent. I realize that you state $x$ and $y$ are complex numbers, but it is in the text and could be missed by the careless reader (i.e. me). However, I am not married to this notation and could easily be convinced just to leave it as you had it.}
Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy integral by
first finding the boundary data of $v$.  Assuming that $\nick{\CC(\Omega)}$ is a
simply-connected interior domain, the boundary data of $v$ satisfies the
Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \nick{\CC(\bd\Omega)}.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Since $v$ is a holomorphic function and we
have its boundary data, by the Cauchy integral theorem we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \nick{\CC(\Omega)}$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation for the Gradient of the
Double-Layer Potential}
\label{sec:gradDLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Computing the shear stress and vorticity requires a complex-valued layer
potential representation of the velocity gradient.  The deformation
tensor at $x \in \Omega$ is found by computing the derivatives of the
expressions for $u_1$ and $u_2$ in equation~\eqref{eqn:cauchyVelocity}  
\begin{equation}
\label{eqn:cauchyGradient}
  \begin{aligned}
    \pderiv{u_1}{x_1} &= +\Re (v'[\psi_1](x)) + 
      \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
      x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
    \pderiv{u_1}{x_2} &= - \Im (v'[\psi_1](x)) - 
      \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
      \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_1} &= +\Re (v'[\psi_2](x)) - 
      \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
      x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_2} &= -\Im (v'[\psi_2](x)) - 
      \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
      \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
  \end{aligned}
\end{equation}
The same expressions are used to compute the deformation tensor for $x
\in \bd\Omega$, except that the jump
condition~\eqref{eqn:deformationJump} is included.  Finally, to compute
the shear stress, the deformation tensor on $\bd\Omega$ is applied to
the normal and tangent vectors as in equation~\eqref{eqn:shearStress}.
The velocity gradient is also used to compute the vorticity in the fluid
bulk.  For $x \in \nick{\CC(\Omega)}$, the Cauchy integral
representation of the vorticity is
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
\Re (v'[\psi_2](x)) + \Im (v'[\psi_1](x)) 
 + \Re (v'[\eta_2](x))+ \Im (v'[\eta_1](x)), \quad x \in \Omega.
\end{align}

@@

The tortuosity definition depends on how the streamlines are
defined, and Duda et al.~\cite{dud-koz-mat2011} summarize several
options.  
\todo[inline]{What does it mean `how streamlines are defined'? I had thought that the definition of a streamline is unambiguous. Do you mean which streamlines are used? i.e.~If there is a bias towards either shorter or longer streamlines?}
\nick{\st{One choice, that we adopt, uses }} The choice adopted here is to use
streamlines that originate at $S$ and terminate at a parallel cross-section.  We use the
cross-sections $x=-1$ and $x=1$ that are at the start and end of the
porous region.  This choice of averaging allows us to also compute the
tortuosity in terms of an area integral.  Assuming that the flow is
incompressible and not re-entrant, meaning that all streamlines connect
the two cross-sections, the tortuosity in
equation~\eqref{eqn:tortuosity1} is equivalent to~\cite{dud-koz-mat2011}
\begin{align}
  T = \frac{\displaystyle\int_\Omega \|\uu(\yy)\|\, d\yy}
           {\displaystyle\int_\Omega u_1(\yy)\, d\yy},
  \label{eqn:tortuosity2}
\end{align}
where $\Omega$ is the fluid region between the inlet and outlet
cross-sections.  Recirculation zones are possible in viscous
fluids~\cite{hig1985}, but they are very small in the examples we
consider and have a negligible effect on the tortuosity.  Since
equation~\eqref{eqn:tortuosity2} does not require the additional work of
computing particle trajectories at every time step, we use this
definition for the majority of the tortuosity calculations.  However, we
do compare the two definitions for the tortuosity at several porosities
in Section~\ref{sec:results}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pore Throat Size}
\label{sec:throats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
While transport in a porous media depends on the porosity, it also
depends on the placement of the grains.  In particular, the grain
placement affects velocity scales~\cite{ali-par-wei-bre2017},
correlation structures~\cite{leb-ded-dav-bou2007}, contaminant
transport~\cite{knu-car2005},
channelization~\cite{sie-ili-pri-riv-gua2019,berhanu2012shape}, and
more. In addition, the grain configuration is important for pore network
models~\cite{bry-kin-mel1993, bry-mel-cad1993, bij-blu2006} that rely on
descriptions of the pore throat sizes. To characterize the grain
placement, we compute distributions of pore sizes between neighboring
grains.  To define neighboring grains, we form a Delaunay triangulation
using nodes placed at the center of each eroding grain and at a
collection of points around the boundary of the porous region.  Then, we
say that two grains are neighbors if their centers share an edge of the
triangulation~\cite{dea-qua-bir-jua2018}.  Finally, we compute the
minimum distance between all pairs of neighboring grains.  In
Figure~\ref{fig:Eroding100gap}(a), we plot a histogram of the pore size
of 100 eroded bodies.  The black curve is an unnormalized Weibull
distribution with the same mean and variance as the measured data.  In
Figure~\ref{fig:Eroding100gap}(b), we plot illustrate the pore throat
sizes by drawing the shortest lines between neighboring grains.  We do
not consider pores between eroding bodies and the solid wall $\Gamma$,
so some of the grains near the porous region boundary only have two
neighbors.  To analyze the effect of erosion on pore sizes, we use a
single Delaunay triangulation to define the neighbors up until a body
has completely eroded. At this point, a new Delanuay triangulation is
formed.

\todo[inline]{This figure seemed out of place to me because you do not present other graphs yet, only the distribution of pores, and it is before the numerical methods are explained. HOWEVER, I think I see what you are driving towards: I think you want to give the reader a little teaser of the results, so that they are more inclined to make it through the methods section without losing interest (is that more or less right?). So perhaps just a few words to indicate our intent would prevent the reader from getting whiplash; maybe phrases like ``As illustration of the statistical results that will be elaborated in Section blah, we show here....''
}

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a) The distribution of the pore
sizes of an eroded geometry. The black curve is the Weibull distribution
with the same first two moments as the data. (b) We only consider pore
sizes between neighboring grains, where two grains are neighbors if they
share a edge in a Delaunay triangulation using nodes placed at the
center of each eroding grain.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\nick{In this section, we apply our new quadrature method to simulate
erosion of several different configurations of nearly-touching eroding
bodies, and to characterize transport inside eroded geometries.}
\todo[inline]{This opening sentence is awkward and seems like a run-on
sentence.} Each eroding body is initially a circle with center $\cc_i$
and radius $r_i$.  The center and radius are chosen at random, and the
body is accepted if it is contained in $[-1,1] \times [-1,1]$ and is
sufficiently separated from all other bodies.  Since the error of our
quadrature method is bounded independent of the geometry, we consider
bodies that are separated by less than 10\% of an arclength spacing.
The randomized method is repeated until the initial geometry reaches a
desired initial porosity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{100 eroding bodies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As a final example, we consider 100 eroding bodies with an initial
porosity near 50\%.  Snapshots of the configurations and vorticity are
in Figure~\ref{fig:Eroding100vort}.  Again, by resolving near-contacts
with the Barycentric quadrature rule, a stable simulation is possible
with only $N_\iin = 256$ discretization points.

@@

We compute the tortuosity using both the
Lagrangian~\eqref{eqn:tortuosity1} and Eulerian
methods~\eqref{eqn:tortuosity2}. Therefore, we must compute the velocity
at $N_p = 1000$ points along the inlet $x=-1$. These normalized
velocities are plotted in Figure~\ref{fig:Eroding100tort}(a) for the
eroded geometry at porosity $\phi = 63.0\%$ \todo[inline]{Make
consistent with caption in figu 13 and another caption somwhere else in
this section}
(Figure~\ref{fig:Eroding100tort}(c)).  The initial velocity of the
tracers is qualitatively similar to the 20 body example
(Figure~\ref{fig:Eroding20tort}(a)), except with additional maximum and
minimums because of the additional bodies.  Next, in
Figure~\ref{fig:Eroding100tort}(b), we plot the local tortuosity by
finding the length of the streamlines as they pass from $x=-1$ to $x=1$.
Compared to Figure~\ref{fig:Eroding20tort}(b), the local tortuosity is
much more discontinuous.  These discontinuities can be explained by
examining the trajectories of tracers in
Figure~\ref{fig:Eroding100tort}(c).  Here, there are many instances of
nearby streamlines that are deflected apart from one another as they
tend to a stagnation point in the flow, and this results in trajectories
with significantly different lengths.  At this porosity, one of the
tracers travels 25.5\% farther than it would have if the bodies had been
absent, and the average tracer travelled 12\% farther than it would have
if the bodies were absent, so the tortuosity is $1.12$.

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local100}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width = \textwidth]{./figs/100b_t100tracer}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100tort} The tortuosity of a porous geometry
that initially had 100 grains and has eroded to a porosity of 63.0\%.
(a) The $x$-component of the velocity at the inlet, $u_1(-1, y)$,
normalized by its maximum velocity $u_{max}=3.90 \times 10^{-4}$.  Note
that this maximum velocity is about a order of magnitude smaller than
when compared to the 20 body example in Figure~\ref{fig:Eroding20tort}.
(b) The local tortuosity $\tau(y)$ on the cross section $x = -1$. The
spatially averaged tortuosity through the channel is $1.12$. (c) The
trajectories of 200 tracers initialized at $x = -1$.  The
discontinuities in the local tortuosity correspond to nearby streamlines
whose lengths are significantly different because they pass opposite
sides of an eroding body. Compared to Figure~\ref{fig:Eroding20tort},
this example has more small bodies, and this results in many more
discontinuities.}
\end{figure}

@@

Finally, we investigate the effects of erosion on pore sizes.  The
distribution of the pore sizes can be directly related to the
distribution of the velocity, and thus effects the
tortuosity~\cite{den-ica-hid2018} and anomalous
dispersion~\cite{dea-qua-bir-jua2018}. In addition, the pore sizes are
required in network models~\cite{bry-mel-cad1993, bry-kin-mel1993}.  As
described in Section~\ref{sec:throats}, we use a Delaunay triangulation
to define neighboring eroding bodies, and we compute the pore size by
finding the closest distance between all neighboring bodies.  Once all
pore sizes are computed, we analyze their distribution as a function of
the porosity.  In Figure~\ref{fig:Eroding100gap_hist}, we plot
histograms of the pore pore sizes at six porosities during the erosion
process. These types of distributions have been described with Weibull
distributions~\cite{ioa-cha1993}, and we superimpose the Weibull
distribution with the same first two moments as the data.  The
parameters of the distribution, $(k,\lambda)$, are included in the
caption of Figure~\ref{fig:Eroding100gap_hist}.  In
Figure~\ref{fig:Eroding100gap_mean_var}, we plot the mean and variance
of the pore sizes as a function of the porosity.  Interestingly, for
porosities less than $\phi = 85\%$, the mean pore size grows linearly
and the variance remains nearly flat. Since a channelized geometry has
large variance, this indicates that channelization is less prevalent at
low porosities.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In equations~\eqref{eqn:velocityDLP} and~\eqref{eqn:deformationDLP},
$\uu$ and its deformation tensor are written as layer potentials in
$\RR^2$, but the quadrature method we introduce in
Section~\ref{sec:method} requires a complex-valued representation of the
layer potentials. The first step to form a complex representation is to
write the Laplace double-layer potential as the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Re (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \Omega,
  \label{eqn:laplaceComplex}
\end{align}
where $x = x_1 + i x_2,y = y_1 + i y_2 \in \CC$ are the complex
counterparts of $\xx = (x_1,x_2),\yy = (y_1,y_2) \in \RR^2$, and $\eta =
\eta_1 + i \eta_2$ is the complex counterpart of $\eeta =
(\eta_1,\eta_2)$. Therefore, depending on the formulation of the layer
potential, the space $\Omega$ can be interpreted as a subset of $\RR^2$
or the corresponding subset of $\CC$.
Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy integral by
first finding the boundary data of $v$.  Assuming that $\Omega$ is a
simply-connected interior domain, the boundary data of $v$ satisfies the
Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \Omega.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Since $v$ is a holomorphic function and we
have its boundary data, by the Cauchy integral theorem we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \Omega$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation for the Gradient of the
Double-Layer Potential}
\label{sec:gradDLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Computing the shear stress and vorticity requires a complex-valued layer
potential representation of the velocity gradient.  The deformation
tensor at $x \in \Omega$ is found by computing the derivatives of the
expressions for $u_1$ and $u_2$ in equation~\eqref{eqn:cauchyVelocity}  
\begin{equation}
\label{eqn:cauchyGradient}
  \begin{aligned}
    \pderiv{u_1}{x_1} &= +\Re (v'[\psi_1](x)) + 
      \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
      x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
    \pderiv{u_1}{x_2} &= - \Im (v'[\psi_1](x)) - 
      \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
      \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_1} &= +\Re (v'[\psi_2](x)) - 
      \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
      x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_2} &= -\Im (v'[\psi_2](x)) - 
      \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
      \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
  \end{aligned}
\end{equation}
The same expressions are used to compute the deformation tensor for $x
\in \bd\Omega$, except that the jump
condition~\eqref{eqn:deformationJump} is included.  Finally, to compute
the shear stress, the deformation tensor on $\bd\Omega$ is applied to
the normal and tangent vectors as in equation~\eqref{eqn:shearStress}.
The velocity gradient is also used to compute the vorticity in the fluid
bulk.  For $x \in \Omega$, the Cauchy integral representation of the
vorticity is
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
\Re (v'[\psi_2](x)) + \Im (v'[\psi_1](x)) 
 + \Re (v'[\eta_2](x))+ \Im (v'[\eta_1](x)), \quad x \in \Omega.
\end{align}

@@

The tortuosity is defined in terms of the length of streamlines between
a starting and ending point. The starting point is typically along a
cross-section $S$, but there are several options for the ending
point~\cite{dud-koz-mat2011}.  The choice adopted here is to use
streamlines that originate at $S$ and terminate at a parallel
cross-section.  We use the cross-sections $x=-1$ and $x=1$ that are at
the start and end of the porous region.  This choice of averaging allows
us to also compute the tortuosity in terms of an area integral.
Assuming that the flow is incompressible and not re-entrant, meaning
that all streamlines connect the two cross-sections, the tortuosity in
equation~\eqref{eqn:tortuosity1} is equivalent to~\cite{dud-koz-mat2011}
\begin{align}
  T = \frac{\displaystyle\int_\Omega \|\uu(\yy)\|\, d\yy}
           {\displaystyle\int_\Omega u_1(\yy)\, d\yy},
  \label{eqn:tortuosity2}
\end{align}
where $\Omega$ is the fluid region between the inlet and outlet
cross-sections.  Recirculation zones are possible in viscous
fluids~\cite{hig1985}, but they are very small in the examples we
consider and have a negligible effect on the tortuosity.  Since
equation~\eqref{eqn:tortuosity2} does not require the additional work of
computing particle trajectories at every time step, we use this
definition for the majority of the tortuosity calculations.  However, we
do compare the two definitions for the tortuosity at several porosities
in Section~\ref{sec:results}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pore Throat Size}
\label{sec:throats}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/triangulation_100b100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_hist100b100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap} (a) The pore sizes between
neighboring grains. Two grains are said to be neighbors if they share an
edge of a Delanuay configuration that uses nodes placed at the center of
each eroding grain. (b) The distribution of the pore sizes of an eroded
geometry. The black curve is the Weibull distribution with the same
first two moments as the data.}
\end{figure}

While transport in a porous media depends on the porosity, it also
depends on the placement of the grains.  In particular, the grain
placement affects velocity scales~\cite{ali-par-wei-bre2017},
correlation structures~\cite{leb-ded-dav-bou2007}, contaminant
transport~\cite{knu-car2005},
channelization~\cite{sie-ili-pri-riv-gua2019,berhanu2012shape}, and
more. In addition, the grain configuration is important for pore network
models~\cite{bry-kin-mel1993, bry-mel-cad1993, bij-blu2006} that rely on
descriptions of the pore throat sizes. To characterize the grain
placement, we compute distributions of pore sizes between neighboring
grains.  To define neighboring grains, we form a Delaunay triangulation
using nodes placed at the center of each eroding grain and at a
collection of points around the boundary of the porous region.  Then, we
say that two grains are neighbors if their centers share an edge of the
triangulation~\cite{dea-qua-bir-jua2018}. The pores for a eroded
geometry are illustrated in Figure~\ref{fig:Eroding100gap}(a). We do not
consider pores between eroding bodies and the solid wall $\Gamma$, so
some of the grains near the porous region boundary only have two
neighbors. Having defined the pore sizes, we form its distribution and
compare it with the Weibull distribution (black curve) in
Figure~\ref{fig:Eroding100gap}(b) . A more complete statistical analysis
of the pore size distribution is in Section~\ref{sec:Eroding100}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We now present numerical results on dense packings of bodies eroding in
Stokes flow. Moreover, we analyze transport through the eroded
geometries by examining coarse-grained variables.  Each eroding body is
initially a circle with center $\cc_i$ and radius $r_i$.  The center and
radius are chosen at random, and the body is accepted if it is contained
in $[-1,1] \times [-1,1]$ and is sufficiently separated from all other
bodies.  Since the error of our quadrature method is bounded independent
of the geometry, we consider bodies that are separated by less than 10\%
of an arclength spacing.  The randomized method is repeated until the
initial geometry reaches a desired initial porosity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{100 eroding bodies}
\label{sec:Eroding100}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As a final example, we consider 100 eroding bodies with an initial
porosity near 50\%.  Snapshots of the configurations and vorticity are
in Figure~\ref{fig:Eroding100vort}.  Again, by resolving near-contacts
with the Barycentric quadrature rule, a stable simulation is possible
with only $N_\iin = 256$ discretization points.

@@

We compute the tortuosity using both the
Lagrangian~\eqref{eqn:tortuosity1} and Eulerian
methods~\eqref{eqn:tortuosity2}. Therefore, we must compute the velocity
at $N_p = 1000$ points along the inlet $x=-1$. These normalized
velocities are plotted in Figure~\ref{fig:Eroding100tort}(a) for the
eroded geometry at porosity $\phi = 62.98\%$
(Figure~\ref{fig:Eroding100tort}(c)).  The initial velocity of the
tracers is qualitatively similar to the 20 body example
(Figure~\ref{fig:Eroding20tort}(a)), except with additional maximum and
minimums because of the additional bodies.  Next, in
Figure~\ref{fig:Eroding100tort}(b), we plot the local tortuosity by
finding the length of the streamlines as they pass from $x=-1$ to $x=1$.
Compared to Figure~\ref{fig:Eroding20tort}(b), the local tortuosity is
much more discontinuous.  These discontinuities can be explained by
examining the trajectories of tracers in
Figure~\ref{fig:Eroding100tort}(c).  Here, there are many instances of
nearby streamlines that are deflected apart from one another as they
tend to a stagnation point in the flow, and this results in trajectories
with significantly different lengths.  At this porosity, one of the
tracers travels 25.5\% farther than it would have if the bodies had been
absent, and the average tracer travelled 12\% farther than it would have
if the bodies were absent, so the tortuosity is $1.12$.

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local100}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width = \textwidth]{./figs/100b_t100tracer}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100tort} The tortuosity of a porous geometry
that initially had 100 grains and has eroded to a porosity of 62.98\%.
(a) The $x$-component of the velocity at the inlet, $u_1(-1, y)$,
normalized by its maximum velocity $u_{max}=3.90 \times 10^{-4}$.  Note
that this maximum velocity is about a order of magnitude smaller than
when compared to the 20 body example in Figure~\ref{fig:Eroding20tort}.
(b) The local tortuosity $\tau(y)$ on the cross section $x = -1$. The
spatially averaged tortuosity through the channel is $1.12$. (c) The
trajectories of 200 tracers initialized at $x = -1$.  The
discontinuities in the local tortuosity correspond to nearby streamlines
whose lengths are significantly different because they pass opposite
sides of an eroding body. Compared to Figure~\ref{fig:Eroding20tort},
this example has more small bodies, and this results in many more
discontinuities.}
\end{figure}

@@

Finally, we investigate the effects of erosion on pore sizes.  The
distribution of the pore sizes can be directly related to the
distribution of the velocity, and thus effects the
tortuosity~\cite{den-ica-hid2018} and anomalous
dispersion~\cite{dea-qua-bir-jua2018}. In addition, the pore sizes are
required in network models~\cite{bry-mel-cad1993, bry-kin-mel1993}.  As
described in Section~\ref{sec:throats}, we use a Delaunay triangulation
to define neighboring eroding bodies, and we compute the pore size by
finding the closest distance between all neighboring bodies.  Instead of
computing a Delanuay configuration at each time step, which would result
in new definitions for the pores at each time step, we only compute a
new Delaunay triangulation when a grain completely erodes.  Once all
pore sizes are computed, we analyze their distribution as a function of
the porosity.

In Figure~\ref{fig:Eroding100gap_hist}, we plot
histograms of the pore pore sizes at six porosities during the erosion
process. These types of distributions have been described with Weibull
distributions~\cite{ioa-cha1993}, and we superimpose the Weibull
distribution with the same first two moments as the data.  The
parameters of the distribution, $(k,\lambda)$, are included in the
caption of Figure~\ref{fig:Eroding100gap_hist}.  In
Figure~\ref{fig:Eroding100gap_mean_var}, we plot the mean and variance
of the pore sizes as a function of the porosity.  Interestingly, for
porosities less than $\phi = 85\%$, the mean pore size grows linearly
and the variance remains nearly flat. Since a channelized geometry has
large variance, this indicates that channelization is less prevalent at
low porosities.
